---
title: "project"
output: html_document
date: "2022-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#Load the dataset
dataset=read.csv("diabetes.csv")
```

Identify any Missing Data Values and replacing them with the mean values of the attribute

```{r}
dataset$Pregnancies = ifelse(is.na(dataset$Pregnancies), ave(dataset$Pregnancies, FUN = function(x) mean(x, na.rm = TRUE )), dataset$Pregnancies)

dataset$Glucose = ifelse(is.na(dataset$Glucose), ave(dataset$Glucose, FUN = function(x) mean(x, na.rm = TRUE )), dataset$Glucose)

dataset$BloodPressure = ifelse(is.na(dataset$BloodPressure), ave(dataset$BloodPressure, FUN = function(x) mean(x, na.rm = TRUE )), dataset$BloodPressure)

dataset$SkinThickness = ifelse(is.na(dataset$SkinThickness), ave(dataset$SkinThickness, FUN = function(x) mean(x, na.rm = TRUE )), dataset$SkinThickness)

dataset$Insulin = ifelse(is.na(dataset$Insulin), ave(dataset$Insulin, FUN = function(x) mean(x, na.rm = TRUE )), dataset$Insulin)

dataset$BMI = ifelse(is.na(dataset$BMI), ave(dataset$BMI, FUN = function(x) mean(x, na.rm = TRUE )), dataset$BMI)

dataset$Diabetes = ifelse(is.na(dataset$Diabetes), ave(dataset$Diabetes, FUN = function(x) mean(x, na.rm = TRUE )), dataset$Diabetes)

dataset$Age = ifelse(is.na(dataset$Age), ave(dataset$Age, FUN = function(x) mean(x, na.rm = TRUE )), dataset$Age)

dataset$Outcome = ifelse(is.na(dataset$Outcome), 0, dataset$Outcome)
```

The Dataset does not contain any type categorical values thus does not require encoding of the values.


Splitting the Dataset into Train and Test Data

```{r}
library(caTools)
set.seed(1)
split=sample.split(dataset$Outcome,SplitRatio = 0.8)
train_data=subset(dataset,split==TRUE)
test_data=subset(dataset,split==FALSE)
```

```{r}
train_data[0:5,]
```

```{r}
dim(train_data)
dim(test_data)
```


```{r}
library(rpart)
library(rpart.plot)
```

Checking if the outcomes are classified correctly using a table format for both train and test data

```{r}
prop.table(table(train_data$Outcome))
```

```{r}
prop.table(table(test_data$Outcome))
```

This formula is to include the attributes required to construct the tree
```{r}
myFormula = Outcome ~ Pregnancies + Glucose + BloodPressure + SkinThickness + Insulin + BMI + DiabetesPedigreeFunction + Age + Diabetes
```

Creating the tree
```{r}
tree <- rpart(myFormula, data = train_data,method = 'class')
```

Plotting the tree
```{r}
rpart.plot(tree, extra = 106)
```

Function to classify the predicted value as either Diabetes or having No Diabetes
```{r}
diabetes_class <- function(hm, sp) {
  if (sp>hm) {
    return("Diabetes")
  }
  else {
    return("No Diabetes")
  }
}
test_data$Pred = test_data$label
```

Predicting the values for test data
```{r}
test_predict <- predict(tree,test_data)
head(test_predict)
```

Using the predicted values to classify them as diabetes or no diabetes
```{r}
for (i in 1:nrow(test_predict)) {
  msg = diabetes_class(test_predict[i,1], test_predict[i,2])
  test_data$Pred[i]=msg
}
```

```{r}
test_data$Pred
```

```{r}
for (i in 1:nrow(test_predict)) {
  out = test_data$Outcome[i]
  if (out == 0) {
    test_data$Outcome[i] = "No Diabetes"
  }
  else {
    test_data$Outcome[i] = "Diabetes"
  }
}
```

```{r}
test_data$Outcome
```

Making the attributes as factors to create confusion matrix
```{r}
test_data$Pred = as.factor(test_data$Pred)
test_data$Outcome = as.factor(test_data$Outcome)
class(test_data$Pred)
class(test_data$Outcome)
```
Creating confusion matrix and misclassification errors

```{r}
tab <- table(as.factor(test_data$Pred), as.factor(test_data$Outcome))
```

Importing caret package which contains confusion matrix function which displays confusion matrix, accuracy, sensitivity, kappa value, etc

```{r}
library("caret")
caret::confusionMatrix(tab)
```
